Case study validations of automatic bullet matching

### 2018-11-06 index.Rmd
light editing

```{r functions, eval=FALSE}
estimateBeta <- function(x, method = "mm", probs = c(0.25, 0.75)) {
  if (method == "mm") {
    mu <- mean(x, na.rm = TRUE)
    sd <- sd(x, na.rm = TRUE)
    lambda <- mu / sd * (1 - mu) / sd - 1
    alpha <- mu * lambda
    beta <- (1 - mu) * lambda
  }
  if (method == "LB") {
    quantile1 <- list(p = probs[1], x = quantile(probs = probs[1], x))
    quantile2 <- list(p = probs[2], x = quantile(probs = probs[2], x))
    bs <- LearnBayes::beta.select(quantile1, quantile2)
    alpha <- bs[1]
    beta <- bs[2]
  }
  return(c(alpha = alpha, beta = beta))
}
```
# Idea of the paper

In \citet{aoas} an algorithm has been laid out to quantify the similarity between land engraved areas of bullets. 
In this paper we present results of the algorithm on several case studies where ground truth is known.

# Introduction and Background
In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with a bullet from the suspected gun under a comparison microscope. They then visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion. Exact guidelines for this classification vary from lab to lab, e.g.\ some labs will exclude only on the basis of non-matching class characteristics (such as direction of the twist in rifling, land length or number of lands, or type of rifling). In some labs,  CMS (consecutively matching striae) as defined by \citet{biasotti} is used as a measure to quantify the strength of a match.

## 3D Imaging Technology
More explicit characterization of bullet surfaces has been discussed since at least 1958 (\citeauthor{davis1968introduction}), but the technology was not sufficiently advanced for analysis based on 3D measurements to be a viable option. \svp{Use of a scanning electron microscope to quantitatively examine and compare bullet striae was demonstrated in \citet{ComputerIdentificationBullets1978}}. More recently, approaches using 3D measurement data were explicitly described in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}. These approaches utilize 3D surface measurements directly to characterize the topology of land engraved areas (LEAs), rather than using visual or image comparison techniques. In many cases, they also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements. Utilizing the 3D measurements allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks. \svp{In addition, there is significantly more information to be gained from quantitative assessment of peak height and valley depth than from the qualitative assessment of the same information.}

>>>>

```{r functions, eval=FALSE}
estimateBeta <- function(x, method = "mm", probs = c(0.25, 0.75)) {
  if (method == "mm") {
    mu <- mean(x, na.rm = TRUE)
    sd <- sd(x, na.rm = TRUE)
    lambda <- mu / sd * (1 - mu) / sd - 1
    alpha <- mu * lambda
    beta <- (1 - mu) * lambda
  }
  if (method == "LB") {
    quantile1 <- list(p = probs[1], x = quantile(probs = probs[1], x))
    quantile2 <- list(p = probs[2], x = quantile(probs = probs[2], x))
    bs <- LearnBayes::beta.select(quantile1, quantile2)
    alpha <- bs[1]
    beta <- bs[2]
  }
  return(c(alpha = alpha, beta = beta))
}
```

# Introduction and Background
In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with a bullet from the suspected gun under a comparison microscope. XXX this could be a second bullet from the crime scene XXX They then visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion. Exact guidelines for this classification vary from lab to lab, e.g.\ some labs will exclude only on the basis of non-matching class characteristics, such as direction of the twist in rifling, land length or number of lands, or type of rifling. In some labs,  CMS (consecutively matching striae) as defined by \citet{biasotti} is used as a measure to quantify the strength of a match.

## 3D Imaging Technology
More explicit characterization of bullet surfaces has been discussed since at least 1958 (\citeauthor{davis1968introduction}), but at the time technology was not sufficiently advanced to make an analysis based on 3D measurements a viable option. \svp{Use of a scanning electron microscope to quantitatively examine and compare bullet striae was demonstrated in \citet{ComputerIdentificationBullets1978}}. More recently, approaches using 3D measurement data were explicitly described in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}. These approaches utilize 3D surface measurements directly to characterize the topology of land engraved areas (LEAs), rather than using visual or image comparison techniques. In many cases, these papers also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements. Utilizing the 3D measurements allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks. \svp{In addition, there is  more information to be gained from quantitative assessment of peak height and valley depth than from the qualitative assessment of the same information.}

\hh{XXX strictly speaking, a comparison microscope is not a qualitative assessment}
